## Exercise 2: football

```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```

Let's first load a dataset:

```{r, echo = TRUE}
fu <- read_csv(file = here::here("data/football.csv"))
```

### Wrangling and tidying

1. Sometimes, we want to know the year of a game happening instead of
   the exact dates. So let's split the dates into three columns:
   `year`, `mon`, and `day`, and save the result tibble to `fu_separated`.
   To do that, you can use:
      - `separate` to separate the first column `date`
        (hint : use `sep = "[-]+"` in `separate`, you'll learn about
         regular expressions later on.)

   To check your answer:

   The output of `print(fu_separated, n = 5)` is

   ```
   # A tibble: 40,945 x 11
     year  mon   day   home_team away_team home_score away_score
     <chr> <chr> <chr> <chr>     <chr>          <dbl>      <dbl>
   1 1872  11    30    Scotland  England            0          0
   2 1873  03    08    England   Scotland           4          2
   3 1874  03    07    Scotland  England            2          1
   4 1875  03    06    England   Scotland           2          2
   5 1876  03    04    Scotland  England            3          0
   # … with 40,940 more rows, and 4 more variables: tournament <chr>,
   #   city <chr>, country <chr>, neutral <lgl>
   ```
   
   ```{r}
   
   fu_separated <- fu %>% 
   separate( date, into = c("year","mon", "day"), sep = "[-]+")
   
   kable(head(fu_separated), "simple")

   ```






2. Notice that the data is recorded per game instead of per team. Suppose
   that we want to have to study the game data per team with the following
   information: `year`, `mon`, `day`, `home_score`, `away_score`,
   `tournament`, `city`, `country`, `neutral`, `team_type`, `team`.
   To do that, we need to list all the games each team played: thus, each
   game will be recorded twice, once for the home team and once for
   the away team.
   Extract such a tibble from `fu_separated` and assign it to `fu_tidy`.

   To check your answer:

   The output of `print(fu_tidy, n = 5)` is

   ```
   # A tibble: 81,890 x 11
     year  mon   day   home_score away_score tournament city  country
     <chr> <chr> <chr>      <dbl>      <dbl> <chr>      <chr> <chr>
   1 1872  11    30             0          0 Friendly   Glas… Scotla…
   2 1872  11    30             0          0 Friendly   Glas… Scotla…
   3 1873  03    08             4          2 Friendly   Lond… England
   4 1873  03    08             4          2 Friendly   Lond… England
   5 1874  03    07             2          1 Friendly   Glas… Scotla…
   # … with 81,885 more rows, and 3 more variables: neutral <lgl>,
   #   team_type <chr>, team <chr>
   ```
   
   ```{r}
   
   fu_tidy <- fu_separated %>% 
   pivot_longer(c(home_team,away_team), "team_type", values_to ="team" )
   
   kable(head(fu_tidy), "simple")
   
   ```

   


3. Now, notice that the two columns `home_score` and `away_score` are
   impractical: they do not indicate whether they correpond to goals scored
   or conceded. Additionally, an important statistic is the goal difference,
   namely the difference between goals scored and conceded.
   From `fu_tidy`, let's extract a tibble `fu_team` that contains
   `year`, `mon`, `day`, `scored`, `conceded`, `gd`, `team`,
   `city`, `country`.

   To check your answer:

   The output of `print(fu_team, n = 5)` is

   ```
   # A tibble: 81,890 x 9
     year  mon   day   scored conceded    gd team     city    country
     <chr> <chr> <chr>  <dbl>    <dbl> <dbl> <chr>    <chr>   <chr>
   1 1872  11    30         0        0     0 Scotland Glasgow Scotland
   2 1872  11    30         0        0     0 England  Glasgow Scotland
   3 1873  03    08         4        2     2 England  London  England
   4 1873  03    08         2        4    -2 Scotland London  England
   5 1874  03    07         2        1     1 Scotland Glasgow Scotland
   # … with 81,885 more rows
   ```
   
   ```{r}
   
   fu_team <- fu_tidy %>% 
   mutate(scored = ifelse(team_type == "away_team", away_score, home_score), 
   conceded = ifelse(team_type == "home_team",away_score, home_score), 
   gd = scored-conceded) %>% 
   select(year:day, scored:gd, team, city, country)
   
   kable(head(fu_team), "simple")
   
   ```
   

   


   

4. The `city` and `country` column represents where the game was held.
   From `fu_team`, extract a tibble `fu_city` that concatenate the city
   and country information into one column called `place`.

   To check your answer:

   The output of `print(fu_city, n = 5)` is

   ```
   # A tibble: 81,890 x 8
     year  mon   day   scored conceded    gd team     place
     <chr> <chr> <chr>  <dbl>    <dbl> <dbl> <chr>    <chr>
   1 1872  11    30         0        0     0 Scotland Glasgow_Scotland
   2 1872  11    30         0        0     0 England  Glasgow_Scotland
   3 1873  03    08         4        2     2 England  London_England
   4 1873  03    08         2        4    -2 Scotland London_England
   5 1874  03    07         2        1     1 Scotland Glasgow_Scotland
   # … with 81,885 more rows
   ```
   
   ```{r}
   
   fu_city <- fu_team %>% 
   unite(place, city, country)
   
   kable(head(fu_city), "simple")
   
   
   ```

   

5. From `fu_city`, let's extract `fu_england`, a tibble containing
   the yearly average goals scored, goals conceded, and goal differences
   for the English team evolved over the years, sorted by descending
   goal differences, descending goals scored, and ascending goals conceded.

   To check your answer:

   The output of `print(fu_england, n = 5)` is

   ```
   # A tibble: 139 x 4
     year  scored conceded    gd
     <chr>  <dbl>    <dbl> <dbl>
   1 1899    6.33    1      5.33
   2 1908    5.57    0.714  4.86
   3 1893    5.67    1      4.67
   4 1888    5       0.667  4.33
   5 1891    5       1      4
   # … with 134 more rows
   ```
   
   
   ```{r}
   
   fu_england <- fu_city %>% 
   group_by(year) %>% 
   filter(team == "England") %>% 
   summarize(scored = mean(scored),
   conceded = mean(conceded), gd = mean(gd)) %>% 
   arrange( desc(gd),  desc(scored), conceded)
   
   kable(head(fu_england), "simple")
   
   ```
   

6. In the following, we want to look at the number of points scored and
   attendance in various international competitions, but first,
   some tidying that is required, as there is some double accounting
   in `fu`.
   Let's create a tibble `fu_tidy2` that contains columns `date`,
   `tournament`, `team_type`, `team`, `scored`, and `year`, where
   `date` and `tournament` as as in `fu`, and
     - `team_type` is `"home_team"` or `"away_team"`,
     - `team` is the name of the national team considered (i.e., either the
       home or away team for each of the two `home_team` and `away_team`
       columns of `fu`),
     - `scored` is the number of points scored by the team under
       consideration,
     - `year` is a numeric extracted from `date`.

   To check your answer:

   The output of `print(fu_tidy2, n = 5)` is

   ```
   # A tibble: 81,890 x 6
     date       tournament team_type team     scored  year
     <date>     <chr>      <chr>     <chr>     <dbl> <dbl>
   1 1872-11-30 Friendly   home_team Scotland      0  1872
   2 1872-11-30 Friendly   away_team England       0  1872
   3 1873-03-08 Friendly   home_team England       4  1873
   4 1873-03-08 Friendly   away_team Scotland      2  1873
   5 1874-03-07 Friendly   home_team Scotland      2  1874
   # … with 81,885 more rows
   ```
   
   ```{r}
   
   fu_tidy2 <- fu %>% 
   pivot_longer(c(home_team,away_team), "team_type", values_to ="team" ) %>% 
   mutate( year = str_split_fixed(date, "-", 3)[,1], 
   scored = ifelse(team_type == "away_team", away_score, home_score)) %>%                                            select(date,tournament, team_type, team, scored, year)
   
   kable(head(fu_tidy2), "simple")
   
   
   ```
   
  
   
  
### Visualization

7. From `fu_tidy2`, let's now look at the yearly number of goals scored from
   2000 to 2019 for teams England, Germany, France, Belgium, Italy and Spain.
   We'll do that by extracting `fu1`, a tibble obtained by grouping by team
   and year and summing the number of goals for each groups,
   before filtering appropriately and transforming `team` as a factor
   sorted by median goals scored.

   To check your answer:

   The output of `print(fu1, n = 5)` is

   ```
   # A tibble: 120 x 3
     team     year scored
     <fct>   <dbl>  <dbl>
   1 Belgium  2000     16
   2 Belgium  2001     25
   3 Belgium  2002     14
   4 Belgium  2003     16
   5 Belgium  2004      6
   # … with 115 more rows
   ```

   The output of `print(levels(fu1$team)[1:3])` is

   ```
   [1] "Belgium" "Italy"   "France"
   ```
      
   ```{r}
   
   fu1 <- fu_tidy2 %>% 
   group_by(team, year) %>% 
   summarize(scored = sum(scored)) %>%
   filter(team %in% c("England","Germany","France","Belgium","Italy","Spain"),
   year >= 2000) %>%
   ungroup(team) %>%
   mutate( team = fct_reorder(team, scored, median)) 
   
   kable(head(fu1), "simple")
   
   ```
   

8. Use `fu1` to produce boxplots of the number of goals scored for each team.
   
   ```{r}
 
   ggplot(fu1, aes( team, scored)) +
   geom_boxplot()
   
   ```

9. Before proceeding to the next plot, let's first create a tibble `fu2`
   containing the relevant data: it should be obtained from `fu_tidy2` by
   transforming `tournament` into a factor containing the levels
   `"FIFA World Cup"`, `"UEFA Euro"`, and `"UEFA Nations League"`,
   dropping NAs (i.e., observations for all other tournaments), and
   filtering for the same 6 teams as well as observations from 2000 onwards.

   To check your answer:

   The output of `print(fu2, n = 5)` is

   ```
   # A tibble: 279 x 6
     date       tournament team_type team    scored  year
     <date>     <fct>      <chr>     <chr>    <dbl> <dbl>
   1 2000-06-10 UEFA Euro  home_team Belgium      2  2000
   2 2000-06-11 UEFA Euro  home_team France       3  2000
   3 2000-06-11 UEFA Euro  away_team Italy        2  2000
   4 2000-06-12 UEFA Euro  home_team Germany      1  2000
   5 2000-06-12 UEFA Euro  away_team England      2  2000
   # … with 274 more rows
   ```

   The output of `print(levels(fu2$tournament))` is

   ```
   [1] "FIFA World Cup"      "UEFA Euro"           "UEFA Nations League"
   ```
   
   ```{r}
   
   fu2 <- fu_tidy2 %>%
   mutate( tournament = factor(tournament, c("FIFA World Cup", "UEFA Euro", "UEFA Nations League"))) %>%
   filter(team %in% c("England","Germany","France","Belgium","Italy","Spain"),
   year >= 2000,
   !(is.na(tournament)))
   
   kable(head(fu2), "simple")
  
   ```
   
print(fu2, n = 5)

print(levels(fu2$tournament))
   

10. Using `fu2`, let's now draw a barplot representing the number of games that
   team England, Germany, France, Belgium, Italy and Spain
   participated in the FIFA World Cup, UEFA Euro, and UEFA Nations League,
   since 2000.
   In the bar plot, each group of columns should represent a team
   and different column for different tournament should be
   shown in different colors.

   ```{r}
   
   
   ggplot(fu2, aes(team, fill = tournament ))+
   geom_bar( position=position_dodge())
   
   ```

11. Use `fu_tidy2` to extract `fu3`, containing a ranking the national teams
   by World cup attendance filtered for those having attended at least 9
   times.

   To check your answer:

   The output of `print(fu3, n = 5)` is

   ```
   # A tibble: 18 x 2
     team          n
     <chr>     <int>
   1 Brazil       21
   2 Germany      19
   3 Italy        18
   4 Argentina    17
   5 Mexico       16
   # … with 13 more rows
   ```
   
   ```{r}
   
  
   fu3 <- fu_tidy2 %>%
   filter(tournament =="FIFA World Cup") %>%
   group_by(team) %>%
   summarize(n = n_distinct(year)) %>%
   filter(n >= 9) %>%
   arrange(desc(n)) 
   
   kable(head(fu3), "simple")
   
   
   ```

print(fu3, n = 5)



12. Then, use `fu3` to create barplot of the world cup attendance
    (i.e., `n`) per team.
    Hints: use `fct_reorder()` to transform `team` as a factor sorted by 
    attendance, and `stat = "identity"`, and `coord_flip()` to flip the axis 
    so that country names can be shown properly.
   
   ```{r}
   
   ggplot(fu3, aes(fct_reorder(team, n), n))+
   geom_bar(stat = "identity")+
   coord_flip()
   
   
   ```
