## Exercise 4: iris

```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```

Let's convert iris to a tibble:

```{r, echo = TRUE}
iris <- as_tibble(iris)
```

### Wrangling and tidying

1. Use `summary()` to on `iris` to compute summaries for all variables
   assign the result to variable `summary_iris`.
   In the remainder of this exercise, we'll try to reproduce the content of
   this summary table by breaking it down in several steps.

   To check your answer:

   The output of `print(summary_iris)` is

   ```
     Sepal.Length    Sepal.Width     Petal.Length    Petal.Width
    Min.   :4.300   Min.   :2.000   Min.   :1.000   Min.   :0.100
    1st Qu.:5.100   1st Qu.:2.800   1st Qu.:1.600   1st Qu.:0.300
    Median :5.800   Median :3.000   Median :4.350   Median :1.300
    Mean   :5.843   Mean   :3.057   Mean   :3.758   Mean   :1.199
    3rd Qu.:6.400   3rd Qu.:3.300   3rd Qu.:5.100   3rd Qu.:1.800
    Max.   :7.900   Max.   :4.400   Max.   :6.900   Max.   :2.500
          Species
    setosa    :50
    versicolor:50
    virginica :50
   ```
   
   ```{r}
options(digits = 4)
summary_iris <- iris %>% summary(digits = 4)
   ```

2. First, let's compute the column that contains the counts across the
   different species of iris and assign it to `species_count`, which should
   contain two colums `Species` and `num`, representing each specie of iris
   and how many of each specie are found in the dataset.

   To check your answer:

   The output of `print(species_count, n = 1)` is

   ```
   # A tibble: 3 x 2
     Species   num
     <fct>   <int>
   1 setosa     50
   # … with 2 more rows
   ```
   
   ```{r}
species_count <- iris %>% group_by(Species) %>% tally %>% summarise(Species,num = n)
   ```

3. Now let's focus on the numerical variables: for all columns except
   for `Species`, compute their minimum, 25th quantile, median, mean,
   75th quantile and maximum, and assign the result to `summary_stats`.

   To check your answer:

   The output of `print(summary_stats, n = 5)` is

   ```
   # A tibble: 1 x 24
     Sepal.Length_Mi… Sepal.Width_Min. Petal.Length_Mi… Petal.Width_Min.
                <dbl>            <dbl>            <dbl>            <dbl>
   1              4.3                2                1              0.1
   # … with 20 more variables:  Sepal.Length_1st Qu.  <dbl>,
   #    Sepal.Width_1st Qu.  <dbl>,  Petal.Length_1st Qu.  <dbl>,
   #    Petal.Width_1st Qu.  <dbl>, Sepal.Length_Median <dbl>,
   #   Sepal.Width_Median <dbl>, Petal.Length_Median <dbl>,
   #   Petal.Width_Median <dbl>, Sepal.Length_Mean <dbl>,
   #   Sepal.Width_Mean <dbl>, Petal.Length_Mean <dbl>,
   #   Petal.Width_Mean <dbl>,  Sepal.Length_3rd Qu.  <dbl>,
   #    Sepal.Width_3rd Qu.  <dbl>,  Petal.Length_3rd Qu.  <dbl>,
   #    Petal.Width_3rd Qu.  <dbl>, Sepal.Length_Max. <dbl>,
   #   Sepal.Width_Max. <dbl>, Petal.Length_Max. <dbl>,
   #   Petal.Width_Max. <dbl>
   ```
   
   ```{r}
   ## Your code goes here
   ```
 
summary_stats <- iris %>% select(-(Species)) %>% summarise(across(everything(), list(Min=min, Q1=~quantile(., probs = 0.25), Median=median, Mean = mean, Q3=~quantile(., probs = 0.75), Max=max), .names = "{.col}_{.fn}")) %>% select(contains("Min"), contains("Q1"), contains("Median"), contains("Mean"), contains("Q3"),contains("Max"))

View(summary_stats)

4. Let's now tidy this tibble to get a result that is easier to understand
   and closer to the output `summary()`: we want to transform this one-row
   tibble into a longer tibble where each row is a variable
   (i.e., one of `Petal.Length`, `Petal.Width`, `Sepal.Length` and
   `Sepal.Width`), and each column is a summary statistics computed for
   the corresponding variable.
   We'll do that in three steps:
   First, we will transform our one-row tibble into a tibble `summary_long`
   with three columns:
     - `variable`, containing the variable name (e.g. `Sepal.Length`),
     - `stat`, containing the name of the summary statistic computed on
       the `variable` of the same row (e.g. `1st Qu.`),
     - `value` containing the value of the `stat` (e.g. `5.1`).
  Assign the result to tibble `summary_long`.

   To check your answer:

   The output of `print(summary_long, n = 5)` is

   ```
   # A tibble: 24 x 3
     variable     stat    value
     <chr>        <chr>   <dbl>
   1 Sepal.Length Min.      4.3
   2 Sepal.Width  Min.      2
   3 Petal.Length Min.      1
   4 Petal.Width  Min.      0.1
   5 Sepal.Length 1st Qu.   5.1
   # … with 19 more rows
   ```
   
   ```{r}
   ## Your code goes here
   ```

summary_long <- summary_stats %>% pivot_longer(everything()) %>% separate(name, c("variable","stat"), sep = "_")

5. As for the second step, from `summary_long`, extract a tibble
   `summary_stats_tidy` in which each row is a variable
   (i.e., one of `Petal.Length`, `Petal.Width`, `Sepal.Length` and
   `Sepal.Width`) and each column is a summary stat computed
   for each variable (i.e., `Min.`, `1st Qu.`, `Median`, `Mean`,
   `3rd Qu.` and `Max.`).

   To check your answer:

   The output of `print(summary_stats_tidy, n = 5)` is

   ```
   # A tibble: 4 x 7
     variable      Min.  1st Qu.  Median  Mean  3rd Qu.   Max.
     <chr>        <dbl>     <dbl>  <dbl> <dbl>     <dbl> <dbl>
   1 Sepal.Length   4.3       5.1   5.8   5.84       6.4   7.9
   2 Sepal.Width    2         2.8   3     3.06       3.3   4.4
   3 Petal.Length   1         1.6   4.35  3.76       5.1   6.9
   4 Petal.Width    0.1       0.3   1.3   1.20       1.8   2.5
   ```
   
   ```{r}
   ## Your code goes here
   ```
 summary_stats_tidy <- summary_long %>% pivot_wider(names_from = stat, values_from = value) 
View(summary_stats_tidy)

6. We are getting closer! The final step: we need to transpose our tibble to
   something similar to `summary(iris)`. Transpose `summary_stats_tidy` into
   a tibble `summary_stats_transposed`. It should have five columns:
   the first, named `Stat` contains the name of the summary statistic
   and the others (`Petal.Length`, `Petal.Width`, `Sepal.Length` and
   `Sepal.Width`) contain the values for each variable.
   To do that, you can successively use `pivot_longer` and `pivot_wider`.

   To check your answer:

   The output of `print(summary_stats_transposed, n = 5)` is

   ```
   # A tibble: 6 x 5
     Stat    Sepal.Length Sepal.Width Petal.Length Petal.Width
     <chr>          <dbl>       <dbl>        <dbl>       <dbl>
   1 Min.            4.3         2            1           0.1
   2 1st Qu.         5.1         2.8          1.6         0.3
   3 Median          5.8         3            4.35        1.3
   4 Mean            5.84        3.06         3.76        1.20
   5 3rd Qu.         6.4         3.3          5.1         1.8
   # … with 1 more row
   ```
   
   ```{r}
   ## Your code goes here
   ```
summary_stats_transposed <-
  summary_stats_tidy %>% pivot_longer(c(`Min`, `Q1`, `Median`, `Mean`, `Q3`, `Max`), names_to = "Stat") %>% pivot_wider(names_from = variable, values_from = value)

### Visualization

7.  Draw a scatterplot of the petal length against sepal length, colored by 
    species, and add a linear regression line for each group.
   
   ```{r}
   ## Your code goes here
   ```
iris %>% ggplot(aes(Sepal.Length,Petal.Length, color=Species)) + geom_point() +  geom_smooth(method=lm)


8. Now, let's produce a scatterplot of sepal width against sepal length, 
   faceted by species, and add a linear regression line for each group.
   
   ```{r}
   ## Your code goes here
   ```
iris %>% ggplot(aes(Sepal.Length,Petal.Length)) +  geom_point()+facet_wrap(~Species) +  geom_smooth(method=lm)

