# 4 Analysis


```{r, echo = FALSE, message = FALSE}
source(here::here("scripts/setup.R"))
```

## 4.1 Research question 1

- Let's put ourselves in context. Our first research question is as follows:

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

Given the organization of public transportion in the city, can we say that the places with the most crime are those that are the most connected by the transportion network?

</div>

<br><br><br>

As we have seen in the EDA of the data set on crime in San Francisco, most crimes are thefts. We wanted to see if, as in some European cities such as Barcelona or Paris, public transportation was a place which facilitated theft. Since the majority of crimes have very little follow-up, we also presumed that criminal activity is not static. Public transportation might therefore be a vector that could influence the choice of target neighborhoods to rob, while also providing a better means of leaving the scene of the crime quickly.

<br>

```{r}
cor(crime_demostats_2016$`Violent crimes`, crime_demostats_2016$`number of stops`
)
```

<br>

We see a very small correlation. It would be interesting to push the analysis a bit further.

<br><br><br>

If we highlight San Francisco's transportation network by comparing it to the segmentation of neighborhoods according to the number of crimes, we can create this map.


<br>

```{r, fig.width=12.5, fig.align='center'}
crime_demostats_2016geo$geometry %>% ggplot() + geom_sf(aes(fill = crime_demostats_2016geo$`Total crimes`)) + scale_fill_gradient(low = "#56B1F7", high = "#132B43") + labs(title = "San Francisco Map : Public Transport vs Crime", fill = "Number of Crime") + xlab("Longitude") + ylab("Latitude") + geom_point(
    aes(x = lon, y = lat)
    ,
    data = transport
    ,
    na.rm = T
    ,
    size = .07,
    color = "yellow"
  ) + theme_dark()
```

<br>

This is the same blue gradient that we have used earlier on to highlight crimes in a neighborhood. The yellow dots represent the public transportation stops in San Francisco.
The network is quite dense and no neighborhood is overlooked. But it seems that the neighborhoods with the most stops are not the ones with the most crime. There are a certain number of stops in Tenderloin, South of Market and Mission (the dark blue neighborhoods) but the areas with the most stops (the 37.8째N/37.78째N and 122.45째N/122.4째N square) are not the ones with the most crime.

<br><br><br>

```{r, fig.width=12.5, fig.align='center'}
crime_demostats_2016 %>% ggplot(aes(`number of stops`, `Violent crimes`)) + geom_point(color = "white") + ylab("Number of Violent Crime") + xlab("Number of Transport Stops") + labs(title = "Violent Crime vs Number of Stops") + geom_smooth(method = "lm", color = "#69b3a2", fill = "black" ) + geom_segment(aes(
  x = 10,
  y = 2200,
  xend = 20,
  yend = 2350
),
arrow = arrow(length = unit(0.5, "cm"))) + geom_segment(aes(
  x = 50,
  y = 1500,
  xend = 43,
  yend = 1800
), arrow = arrow(length = unit(0.5, "cm"))) + geom_segment(aes(
  x = 70,
  y = 2200,
  xend = 80,
  yend = 2400
), arrow = arrow(length = unit(0.5, "cm")))  + theme_dark()
```

<br>

We can see some small correlation between the number of crimes and the number of bus stops. On the other hand, the 3 neighborhoods with the most crimes, which we have highlighted with 3 arrows, seem to be outliers that are very important for the interpretation of the results. This very small correlation is not very statistically significant given the variability of the observations. We can see that the black area, which represents the standard error, is very large.

<br><br>

```{r}
cor.test(crime_demostats_2016$`Violent crimes`, crime_demostats_2016$`number of stops`)
```

```{r}
busmod <- lm(`Total crimes`~ `number of stops` , data = crime_demostats_2016)
summary(busmod)
```

<br><br>

As expected, this small correlation is not statistically significant (p-value = 0.7995 and 0 is in the 95% confidence interval). In addition, the variation in the number of crimes explained by the variation in the number of stops between neighbourhoods is extremely small. (Multiple R-squared = 0.00167). We can therefore conclude that there is no linear relationship between the number of crimes and the number of public transit stops. We cannot conclude that the majority of flights frome crime scenes do not take place that much when travelling by public transport, which was one of our theories.

<br><br><br>

## 4.2 Research question 2

<br>

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

Here we will explore whether the COVID-19 pandemic had any effect on crime rates.

</div>


We will first compare the total number of crimes citywide in the past 5 years, limiting the time period to the period between March 10th and November 13th so that all numbers are comparable to the start of the pandemic in the United States and the last date of our 2020 crime dataset.


```{r}
crime_2015_2020_year <- crime2003_2020_datesep %>% 
  mutate(month_day = as.numeric(format(`Incident Date`, "%m%d"))) %>%
  filter(year %in% c(2015,2016,2017,2018,2019,2020),  
           month_day > 310 &
           month_day < 1113) %>%
  count(`year`, name = "Total crimes")
kable(crime_2015_2020_year, "simple")
```

<br>

There seems to be a clear drop in crime in 2020 ove the COVID-19 period relative to previous years. The numbers for this approximately 8-month period in the previous years has always been cloe to 10'000 total crimes, while in 2020 we have have less than 7'500, a more than 25% drop.

In this time series graph, where we compare crimes since the beginning of 2018 and COVID-19 case counts, we can clearly see that once the infections line appears on the graph the crime rate line seems to sustainably drop into a lower zone. While both sets of data have some level of short-term cyclical variation (on weekends, we see more crimes , but fewer recorded coronavirus cases), taking this long view of time we can clearly see that something has changed with crime rates in a more durable way since the start of the pandemic.



```{r}
options(highcharter.theme = hc_theme_smpl(tooltip = list(valueDecimals = 2)))
crime_2018_2020_date <- crime2003_2020 %>%
  filter(  `Incident Date` <= as.Date("2020-11-13") &
             `Incident Date` >= as.Date("2018-01-01") ) %>%
  group_by(`Incident Date`) %>%
  count( name = "Total crimes") %>%
  rename(`date` = `Incident Date`)
 
 
 crime_covid_long_2018_2020 <- merge(crime_2018_2020_date, covid_sf, by="date", all = T) %>%
   pivot_longer(c(`Total crimes`, `Case Count`),
                names_to = "variable",
                values_to = "value")
  crime_covid_long_2018_2020 %>% hchart("line", hcaes( `date`, `value`, group = "variable")) 
```

<br>

This graph shows the timeline of crime rates for 2018, 2019 an 2020 and case counts for 2020 over the course of their respective years. We can see clearly that 2020 was following the same trend in crime rates as in the two previous years in January eand February, there is no real distinction between the three lines. Once the pandemic comes into play, the 2020 crime line becomes clearly distinct, at a lower rate than the two lines representing previous years.

```{r}
  crime_covid_long_2018_2020_noyears <-
    crime_covid_long_2018_2020 %>% mutate(
      year = as.character(format(`date`, "%Y")),
      date = (format(`date`, "%m-%d")),
      variable = paste(variable, year)
    ) %>% select(date, variable, value) %>% filter(!(date == "02-29") &
                                                     !(variable %in% c("Case Count 2018", "Case Count 2019")))
  
  crime_covid_long_2018_2020_noyears %>% hchart("line", hcaes(`date`, `value`, group = "variable"))
```

<br>

While it would be important to look into the exact timeline of lockdown measures, it would seem that the decrease in interactions since the beginning of the pandemic, through government measures and individual choices, has led to a decrease in reported crime rates.

<br><br><br>

## 4.3 Research question 3

<br>

<style>
div.blue { background-color:#e6f0ff; border-radius: 5px; padding: 20px;}
</style>
<div class = "blue">

**Does the crime rate vary significantly by neighborhood? Do factors such as education, income, ethnicity of the population, poverty, unemployment rate, population density explain this variation?**
We will try to predict the number of crimes in a neighbourhood with it's socio-economic characteristics. Do these factors tell us something about the perpetrators of crimes and their victims ? We will try to see if there is any linear relationship.
</div>

<br><br><br>


#### Variable Selection

```{r}
par(mfrow = c(1, 1))
corrplot::corrplot(cor(crime_demostats_2014_2016), method = "shade",tl.pos = 'n')
```

<br>

To begin with, the data set we're going to use has a lot of variables. Instead of making a step-wise selection on the total number of variables, we will first analyze which independent variable is most correlated with our dependent variable. Using the first line of this correlation matrix, we selected the 6th, 13th, 27th, 34th and 38th columns of our data set. We will try to explain the total number of crimes with the variables `Non-Family Households`, `Other/Two or More Races`, `Units of Housing`, `Median Household Income` and `Number of people Poverty`.

<br><br><br>

#### Model 1

<br>

We make our first linear model:

$$
\begin{align}
    Total Crimes = Non-Family Households + Poverty + Other/Two More Races + Housing+ Income
\end{align}
$$

```{r}
fit2.1 <- lm(
  `Total crimes` ~`Number of people Poverty` + `Other/Two or More Races` + `Units of Housing` + `Non-Family Households`+ `Median Household Income`, data = crime_demostats_2014_2016
)
summary(fit2.1)
```

<br><br>

Not all of our variables are significant. To have a better model, we will use a backward stepwise selection so that all our independent variables are statistically significant. The least significant variable is the one related to the ethnicity of the population. To simplify our model, we will remove this variable.

<br><br>

#### Model 1.2

<br>

$$
\begin{align}
    Total Crimes = Non-Family Households + Poverty + Housing+ Income
\end{align}
$$

```{r}
fit2.1_2 <- lm(
  `Total crimes` ~`Number of people Poverty` + `Units of Housing` + `Non-Family Households`+ `Median Household Income`, data = crime_demostats_2014_2016
)
summary(fit2.1_2)
```

<br>

This time, the least significant variable is the wage variable. We remove this variable.

<br><br>

#### Model 1.3

<br>

$$
\begin{align}
    Total Crimes = Non-Family Households + Poverty + Housing
\end{align}
$$

```{r}
fit2.1_3 <- lm(
  `Total crimes` ~`Number of people Poverty` + `Units of Housing` + `Non-Family Households`, data = crime_demostats_2014_2016
)
summary(fit2.1_3)
```

<br>

We have chosen to select a significance level of 5%. As the variable `Units of Housing` has a p-value that is greater than 5%, we will also remove it. We remove this variable..

<br><br>

#### Model 1.4

<br>

$$
\begin{align}
    log(Total Crimes) = log(Non-Family Households) + Poverty^2
\end{align}
$$

```{r}
fit2.2 <- lm(
  log(crime_demostats_2014_2016$`Total crimes`) ~ log(`Non-Family Households`) + I(`Number of people Poverty`^2), data = crime_demostats_2014_2016)
summary(fit2.2)
```
<br>

To reduce the effects of non-normality of the distribution, we chose to perform some transformations on the variables. All of our variables are statistically significant. Moreover, the Multiple R-squared is not so bad. It seems to be a good model. 

```{r}
car::vif(fit2.2)
```
We also see that the correlation between the variables is not high at all. For the moment, all the conditions for a linear regression seem to be met.

<br>

We still need to check the error distribution to see if our conclusions can be used.
```{r, fig.width=10, fig.height=2, fig.fullwidth=TRUE}
plot(fit2.2)
```
<br>

As can be seen on the normal QQ graph, our errors are not normally distributed at the extremes. On the Residuals vs Fitted graph, we see a certain pattern, which would mean that our errors are not constant. With this graph, we also see outliers: observations 8, 49 and 90. These three observations represent the same neighborhood of San Francisco at three different time intervals. However, this neighborhood is a park with very few inhabitants, making the demogrpahic data on this place not very relevant. We have chosen to delete these three observations.

<br><br><br>

```{r}
crime_demostats_2014_2016_2 <-
  crime_demostats_2014_2016[c(-8,-49, -90), ] #remove the outliers
fit2.3 <- lm(
  log(`Total crimes`) ~ log(`Non-Family Households`) + I(`Number of people Poverty`^2), data = crime_demostats_2014_2016_2)
summary(fit2.3)
```

```{r}
car::vif(fit2.2)
```

<br>

Very little change in the significance and explanation of the variation of the dependent variable by the independent variables by removing the three observations considered as outliers.

<br>

Let's take a close look now to see if the assumptions of normality on errors seem better:
```{r, fig.width=10, fig.height=2, fig.fullwidth=TRUE}
plot(fit2.3)
```

Consistency of errors seem to be respected. However, the extreme values still deviate a little from the values we should have. The distribution of errors does not totally follow a normal law but it is close enough for our model be relevant.

<br><br><br>


What can we conclude from this model?
Since this kind of model is just an explanation of correlation, we cannot establish a causal link between the independent variables and the dependent variable. Moreover, like all social studies, there is a lot noise that make it difficult to explain the phenomenon that we are studying.
But we will nevertheless propose various hypotheses that can explain our results.

<br>

First, the two most predictive variables for crime by neighborhood are poverty and the presence of non-family households. This could be reflective of the nature of the nature of the perpetrators, victims or just the neighborhood in itself. These neighbourhoods probably have a lot of students and young professionals, indicating above average rates of social activity and nightlife.  Criminals might choose such neighborhoods to commit crimes such as theft  because it is easier to pass unnoticed in active spaces. To confirm this hypothesis, it would have been interesting to add a database of entertainment locations in San Francisco, to have the description of the victims and information on the density of social interaction by neighborhood. 
It is also possible that this indicates the demographic categories who are more likely to commit crimes. Risky behaviour such as criminal activity might be more common among people who are poorer and do not have families.

```{r, fig.width=12.5, fig.height=2, fig.fullwidth=TRUE, out.width = '100%', fig.align='center'}
c1 <- demostats_2016geo$geometry %>% ggplot() + geom_sf(aes(fill = demostats_2016geo$`Non-Family Households`)) + scale_fill_gradient(low = "#56B1F7", high = "#132B43") + xlab("Longitude") + ylab("Latitude") + labs(title = "Unstable situation neighborhood", fill = "Percentage of Non-Family Households") 

c2 <- demostats_2016geo$geometry %>% ggplot() + geom_sf(aes(fill = demostats_2016geo$`18-34 years`)) + scale_fill_gradient(low = "#56B1F7", high = "#132B43") + xlab("Longitude") + ylab("Latitude") + labs(title = "Young Neighborhood", fill = "Percentage of of 18-34 years people") 

c3 <- demostats_2016geo$geometry %>% ggplot() + geom_sf(aes(fill = demostats_2016geo$`Percent in Poverty`)) + scale_fill_gradient(low = "#56B1F7", high = "#132B43") + xlab("Longitude") + ylab("Latitude") + labs(title = "Poverty", fill = "Percentage of Poor in the neighborhood")

c4 <- demostats_2016geo$geometry %>% ggplot() + geom_sf(aes(fill = demostats_2016geo$`Median Household Income`)) + scale_fill_gradient(low = "#56B1F7", high = "#132B43") + xlab("Longitude") + ylab("Latitude") + labs(title = "Income: Look the lowest average income neighborhood ", fill = "Average Income in $")

c1 + c2 + c3 + c4

```


In an attempt to add weight to our hypothesis, we wanted to make a geographical representation of factors that represent characteristics of young people: poverty, presence of non-family households situation, an age range between 18 and 35 years old and a low salary.
Looking at that these 4 variables, the neighbourhoods with the most crime are often linked to these characteristics, but not always in the same way. To take an example, Mission stands out as
It would be interesting to carry out further research !

The second proposal of interpretation that we had thought of is more in the sense of an analysis of an explanation of the crime.
Let's take the case of static crime (the one that stays in one's neighborhood), we could explain the fact that people steal with the fact that they are poor and that they are in an unstable situation. Theft would be a solution that would alleviate their worries.